---
title: 'Philly Center City Sips 2019: An Interactive Map'
author: silvia
date: '2019-08-04'
slug: ccd-sips
categories: [R minis]
tags:
  - mapping
  - R
subtitle: 'An interactive map showing restaurants participating in Philly’s Center City District Sips 2019.'
summary: 'An interactive map showing restaurants participating in Philly’s Center City District Sips 2019.'
authors: []
lastmod: ''
featured: yes
image:
  caption: '[Image credit: Silvia Canelón](featured.png)'
  placement: 2
  focal_point: 'Center'
  preview_only: no
projects: []
links:
- icon: map-marked-alt
  icon_pack: fas
  name: Interactive Map
  url: https://spcanelon.github.io/CCDSips2019.html
---

I used **R** to play around with **webscraping** and create an [interactive map](https://spcanelon.github.io/CCDSips2019.html/) showing restaurants participating in Philly’s 2019 [Center City District Sips](https://centercityphila.org/explore-center-city/ccdsips). Center City Sips is a series of summer Wednesday evenings filled with **happy hour specials**.

This mini-project was motivated by two of my close friends' frustration with the website shortcomings. **The information was presented beautifully, but there was no map view**! This made it difficult to figure out what happy hours you could check out nearby.

The map pop-ups could use information on the specials for each restaurant, but my newly acquired skills with the `RSelenium` and `Leaflet` packages weren't quite up to the task!

<div style="align: center; margin-left: -150px;"> <iframe src="https://spcanelon.github.io/CCDSips2019.html" width="1000px" height="900px" frameborder="0"></iframe> </div> 

```{r setup, eval=TRUE, include=FALSE}
library(tidyverse)
library(ggmap)
library(leaflet)

# importing data
sips <- readRDS("sips.rds")
```

```{r create-map, eval=FALSE, echo=FALSE, out.width=2000}
# Create the pop-up information with inline css styling
popInfoCircles<-paste("<h2 style='border-bottom: thin dotted #43464C;
    padding-bottom:4px; margin-bottom:4px;
    font-family: Tahoma, Geneva, sans-serif;
    color:#43464C;'>", sips$Restaurant, "</h2>
    <h3 style=font-family: Tahoma, Geneva, sans-serif;
    color:#9197A6;'>", sips$Address, "</h3>")

popInfoMarker<-paste("<h1 style='border-bottom: thin dotted #43464C;
    padding-bottom:4px; margin-bottom:4px;
    font-family: Tahoma, Geneva, sans-serif;
    color:#43464C;'>", "Center City District Sips 2019", "</h1>
    <span style='color:#9197A6;'>", "Philadelphia, PA", "<br>")

# Creating the map
leaflet(data = sips) %>%
  addCircles(lat = ~ sips$Latitude, lng = ~ sips$Longitude, color = "olivedrab",
             fillOpacity=0.6, radius=8, popup = popInfoCircles) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(mean(sips$Longitude), mean(sips$Latitude), zoom=17) %>%
  addMarkers(lng = mean(sips$Longitude), lat = mean(sips$Latitude), popup = popInfoMarker)
```
