---
title: 'Migrating from Trello to Airtable: Working with JSON Data in R'
author: silvia
date: '2020-05-10'
slug: trello-to-airtable
categories: []
tags:
  - r
  - json
subtitle: 'Tutorial on how to migrate your current Trello board to an Airtable base from start to finish.'
summary: 'Tutorial on how to migrate your current Trello board to an Airtable base from start to finish.'
authors: []
lastmod: '2020-05-09T23:19:13-04:00'
featured: yes
draft: no
image:
  caption: '[Trello to Airtable migration using the example in this tutorial](featured.png)'
  focal_point: 'Center'
  placement: '2'
  preview_only: no
projects: []
output:
  blogdown::html_page:
    toc: true
    number_sections: true
    toc_depth: 1
links:
  - icon:
    icon_pack:
    name: Rmd
    url: 2020-05-10-trello-to-airtable.Rmd
---

**[Airtable](https://airtable.com/) is a user-friendly and _powerful_ tool** that until recently I'd been using for personal projects (i.e. document organizing, apartment hunting, etc.). A week ago I leaned on Airtable to create a base designed for the [Philadelphia Reproductive Freedom Collective](https://www.womensmedicalfund.org/prfc) to support our COVID-19 mutual aid efforts. 

> Shout out to the [Silicon Valley Democratic Socialists of America](https://siliconvalleydsa.org/) who provided inspiration and support! Their mutual aid network is extensive and operates at an impressive scale. Get connected with the South Bay Mutual Aid network here: https://www.aidnetwork.org/

Having fallen in some Airtable deep-work I figured maybe it was time to **retire my Trello boards** in favor of some task bases. Airtable accepts CSV files from Trello but, alas, my [_free_ Trello account](https://trello.com/pricing) only gave me the option to print as a PDF or export as JSON. I decided this would be a good opportunity to learn how to parse JSON data and export it as a CSV ready for import to Airtable.

# What to expect
The code provided in the [R Markdown file](https://github.com/spcanelon/silvia/tree/master/content/post/20200505-trello-to-airtable-example) requires the `tidyverse`, `JSONlite`, and `lubridate` packages as well as a JSON file containing your Trello data. It will generate the following:

- `tidy_cards.csv` file ready to import to Airtable 
- `attachments` folder containing any content you've attached to your cards using an URL. 
- `attachment_errors.csv` file containing the card name, or task, as well as the URL/file path of any attachments that failed to download (i.e. links to files stored locally), so that you exactly which links you'll need to locate manually.

**By the end of the tutorial** you'll have imported a CSV file into Airtable _**and**_ configured Airtable to match your Trello board. Specifically, you'll be able to migrate labels, lists, attachments, card names, card descriptions, card due dates, date last modified, and whether the cards were marked "complete" or archived.

## What _not_ to expect
- Rendered markdown from your card descriptions. Airtable _does_ support markdown in its "long text" field type, but it doesn't automatically render from imported text.

# Choose your own adventure
1. **Download the R Markdown file** linked above <i class="fas fa-hand-point-up"></i> and run with it. Jump down to section [Formatting in Airtable](#formatting-in-airtable) for pointers on how to get your Airtable base configured like your Trello board.
2. **Keep reading** for a code-through of the R Markdown file, using the example featured in the picture above, with tips and resources along the way.

# Follow along with me!

The Trello board used for this code-through is this [Program Management Template](https://trello.com/templates/project-management/program-management-template-potjK48u) created by Soniya Ahuja. To download the Trello board data, go to _Menu_ > _More_ > _Print and Export_ and then right-click on "Export as JSON" to save the JSON file in your working directory.

You may also choose to read through Trello's [Making Sense of Trello's JSON Export](https://help.trello.com/article/924-making-sense-of-trellos-json-export).

Last but not least -- we'll be keeping things tidy, with the help of the [**tidyverse** collection of R packages](https://www.tidyverse.org/), to ensure each variable is in its own column, observations are in rows, and values are in cells. You can read more about tidy data in [Chapter 12 of R for Data Science](https://tidyr.tidyverse.org/articles/tidy-data.html).

# Importing JSON file

```{r, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r import, results=FALSE}
# clearing environment
rm(list = ls())

# loading libraries
library(jsonlite)
library(lubridate)
library(tidyverse)
```

```{r, include=FALSE}
library(emo)
```

```{r, results=FALSE}
trello <- stream_in(file("program-mgmt.json")) # produces a data frame
```

```{r}
glimpse(trello)
```

# Parsing information

Instructions found here: https://blog.exploratory.io/working-with-json-data-in-very-simple-way-ad7ebcc0bb89

## Cards
- *Flattening:* Making the nested hierarchical data structure into a flatter structure by assigning each of the nested variable as its own column as much as possible.
- *Tibble:* Making it easier to see the data frame data

```{r tibble-time, paged.print=TRUE}
# selecting cards information
cards <- trello$cards # list of 1

# flattening
cards_flat <- flatten(cards) #list of 37

# tibble time
cards_flat_tbl <- as_tibble(cards_flat) # 32 obs of 37 variables
glimpse(cards_flat_tbl)

# removing unwanted variables
cards_trim <- cards_flat_tbl %>%
  select(-address, -checkItemStates, -coordinates, -creationMethod, -descData, -dueReminder,
         -idAttachmentCover, -idBoard, -idLabels, -idMembersVoted, -limits, -locationName, -manualCoverAttachment,
         -(pos:badges), -(email:idMembers), -subscribed, -url, -cover, -pluginData, -customFieldItems) %>%
  arrange(desc(dateLastActivity)) %>%
  select(id, idShort, idList, dateLastActivity, name, desc, dueComplete, due,
         labels, attachments, shortUrl, closed)
```

### Labels

Resources:
- **Used:** the `unnest` function after selecting the minimum necessary variables
- Helpful, again! https://blog.exploratory.io/working-with-json-data-in-very-simple-way-ad7ebcc0bb89 , particularly for saving the label details as a character list

```{r labels}
# extracting labels details
labels_info <- cards_trim %>%
  select(id, idShort, labels) %>%
  unnest() %>% # no arguments because the nested items don't have names
  rename(labelName = name) %>%
  select(id, idShort, labelName) %>%
  group_by(id, idShort) %>%
  summarize(labelList = list(labelName)) %>%
  mutate(labelList = as.character(labelList)) %>%
  mutate(labelList_tidy = str_remove_all(labelList, pattern = "\"")) %>%
  mutate(labelList_tidy = str_remove_all(labelList_tidy, pattern ="c\\(")) %>%
  mutate(labelList_tidy = str_remove_all(labelList_tidy, pattern ="\\)")) %>%
  unique()

knitr::kable(labels_info %>% head(n = 3L))

# joining back with main cards data frame
ct_labels <- left_join(cards_trim %>% select(-labels), labels_info %>% select(-labelList))
```

### Attachments

The next step is to download all of the items attached the cards onto a local folder. I found [this StackOverflow post](https://stackoverflow.com/questions/32174306/download-url-links-using-r) really helpful. When I tried this out on my own Trello board I also found that I couldn't download the few attachments I had made from my local drive. [This StackOverflow post](https://stackoverflow.com/questions/2158780/catching-an-error-and-then-branching-logic/2158803#2158803) helped me figure out how to flag and catch these download errors so that I could create a list of the urls with "attachment errors" that I could follow up with manually. 

```{r}
# expanding the attachment lists into separate url records
att_urls <- ct_labels %>%
  select(idShort, attachments) %>%
  unnest() %>%
  select(idShort, url) %>%
  mutate(url = as.character(url),
         attachmentError = 'FALSE')
knitr::kable(att_urls %>% head(n = 3L))

# create directory for attachments
dirAttachments <- "attachments/"
dir.create(dirAttachments)

# downloading urls and checking for errors using try()
for (i in 1:length(att_urls$url)){
  locAttachments <-
        paste(dirAttachments, "/", att_urls$idShort[i], "_", basename(att_urls$url[i]), sep = "")
  step_to_try <- try(attachment_check <- download.file(att_urls$url[i], destfile = locAttachments))
  if("try-error" %in% class(step_to_try)) {
    cat("Error row: ", i, "\n", "Error message: ", step_to_try[1], sep = "")
    att_urls$attachmentError[i] = 'TRUE'
  }
}
```

**Side step:** If you have a lot of attachments _per card_, you may want to create a directory folder for each card. This for loop will get you there -- use it instead of the one above.
```{r, eval=FALSE}
# creates individual directory folders for each card id -- probably overkill and more cumbersome for import purposes
for (i in 1:length(att_urls$url)){
  dirAttachments <- paste(dirFiles, "attachments", att_urls$idShort[i], sep = "/")
  dir.create(dirAttachments) # creates directory for each unique card id
  locAttachments <- paste(dirAttachments, basename(url[i]), sep = "/")
  download.file(url[i], destfile = locAttachments)
}
```

```{r}
# preparing data frame for export to CSV
attachment_errors <- att_urls %>%
  filter(attachmentError == TRUE) %>%
  rename(Task_Id = idShort, Attachment_URL = url, Attachment_Error = attachmentError)

# converts the attachment column to a categorical variable in the main cards data frame
ct_labels <- ct_labels %>%
  mutate(attachments = ifelse(idShort %in% att_urls$idShort, TRUE, FALSE))

# exporting to CSV
write.csv(attachment_errors, file = "attachment_errors.csv")
```

## Lists
Transforming the "lists" list to a data frame was easy because there was only one data frame in it. If there had been more, I would have used the `unnest` function as before
```{r}
# selecting lists information
lists <- trello$lists # list of 1 data frame
glimpse(lists)

# flattening
lists_flat <- lists[[1]] # 17 obs of 9 variables

# removing unwanted variables
lists_trim <- lists_flat %>%
  select(id, name, closed) %>%
  rename(idList = id, nameList = name, closedList = closed)
knitr::kable(lists_trim %>% head(n = 3L))

# joining back with main cards data frame
ct_labels_list <- left_join(ct_labels, lists_trim) %>%
  select(id:shortUrl, labelList_tidy:nameList, closed, closedList)
```

# Exporting to CSV

## Data prepping
The `lubridate` package was used next to convert the date fields and [this resource](http://biostat.mc.vanderbilt.edu/wiki/pub/Main/ColeBeck/datestimes.pdf) was helpful in understanding date conversions and formatting.
```{r export-prep, paged.print=TRUE}
# changing variable names
tidy_cards <- ct_labels_list %>%
  select(-id, -idList, -closedList) %>%
  rename(Task = name, Task_ID = idShort, Notes = desc, Done = dueComplete, Date_Due = due,
         Labels = labelList_tidy, Trello_List = nameList, Trello_Last_Modified = dateLastActivity,
         Trello_Url = shortUrl, Trello_Attachments = attachments, Archived = closed) %>%
  select(Task, Task_ID, Notes, Done, Date_Due, Labels, Trello_List, Trello_Last_Modified,
         Trello_Url, Trello_Attachments, Archived) %>%
  mutate(Trello_Last_Modified = as_datetime(Trello_Last_Modified, tz = ""),
         Date_Due = as_datetime(Date_Due, tz = ""),
         Done = ifelse(is.na(Date_Due) == TRUE, 'NA', Done)) # ensures only undone tasks assigned a due date get marked as "FALSE"
```

```{r collapse=TRUE}
length(unique(tidy_cards$Task_ID)) # confirm this matches the number of records (i.e. one task per observation)

glimpse(tidy_cards)
```

## Data exporting
```{r export-csv, eval=TRUE}
write.csv(tidy_cards, file = "tidy_cards.csv")
```

# Formatting in Airtable
Import CSV into Airtable as a new base

- **Add a base** > **Import a spreadsheet** > **CSV file**. Upload the `cards_tidy.csv` file. Pick a name and icon for your new base! `r emo::ji('tada')` More detailed instructions can be found in [this Airtable article](https://support.airtable.com/hc/en-us/articles/202579399-Import-an-Existing-Spreadsheet-or-CSV-Into-a-New-App#cleanup). 

## Task
1. Because we know each observation in our table is unique, we can copy and paste `Task` into the first column and hide/delete the original column. `Task` is now the primary field.

## Task_ID
1. Convert `Task_ID` field type to "number".

## Notes
1. Convert `Notes` field type to "long text" and enable _rich text formatting_. This gives us the option of using Markdown in the future, but sadly doesn't automatically recognize fully formatted Markdown in the imported text `r emo::ji('disappointed')`.

## Labels
1. Change `Labels` field type to "multiple select" so that it turns each item in each list into a label.
2. _**Optional**_: Create a new `Projects` column next to `Labels` and use the labels to guide you in creating `Project` labels/categories: Group the records by `Labels` field and add to `Projects` field as appropriate.

{{% alert note %}}
I recommend creating an `NA` project from the `NA` labels so that these tasks aren't marked as "uncategorized" in the `Projects` column. Having records with an "empty" assignment gets in the way whenever you want to group by that category. To that end, it's helpful to group by `Project` and make sure any "empty" records get assigned to the "NA" `Project`.
{{% /alert %}}

3. Delete from `Labels` any labels that were converted to `Projects` and ungroup the records.

## Trello_Lists
1. Convert `Trello_Lists` column field type into "single select". This gives us the option of using the Kanban style we were used to in Trello.

{{% alert note %}}
Every record should already be associated with a Trello_List.
{{ /alert %}}

2. If you want to replicate the Trello kanban layout, change the order of the single select options in `Trello_Lists` to match the order from left-to-right of your Trello board

## Trello_URL
1. Convert the `Trello_URL` field type to "URL" and then hide it if you don't think you'll reference it often.

## Trello_Attachments
1. Convert `Trello_Attachments` field type to "single select"
2. Create a new `Attachments` with field type "attachment". This is where you'll upload your downloaded attachments.
3. Filter your records by `Trello_Attachments` so only show "TRUE" results
4. Sort records by `Task_ID` and simplify your view by temporarily hiding all columns except for `Task_Name` (primary field), `Task_ID`, `Trello_Attachments`, and the new `Attachments` column.
5. Open your local `attachments` folder and drag and drop the files to their corresponding `Attachments` field according to their `Task_ID` in the filename.

{{% alert note %}}
Increase the height of the records for this step. It'll make it easier to make sure you're dragging and dropping to the correct record
{{ /alert %}}

6. If you encountered errors downloading some of your attachment URLs, now is the time to check your local _attachment_errors.csv_ file for the URLs with errors during the download process. These are attachments you'll have to find elsewhere and upload to the `Attachments` field as needed.
7. Remove the filter to your view and unhide any columns you wish to remain visible.

## Done & Archived
1. Convert the `Done` and `Archived` field types to "Checkbox" and it will automatically assign a "check" to all records marked "TRUE" and leave the ones marked "FALSE" or "NA" unchecked. So easy!
2. There is no direct option to "archive" tasks that have been completed like you can do in Trello, _but_ you can apply a filter to your table view to hide the tasks that are complete. To do this, set the filter so that the `Done` and `Archived` fields are unchecked.

  > **Note:** This must be repeated for each saved View of your records.

# Moving forward

## Dates
1. Modify `Date_Due` and `Trello_Last_Modified` field types to "Date" with time.
2. You can sort the records by `Trello_Last_Modified` if that's helpful, but otherwise you can hide the column and keep it for historical reference.
3. Create a new column `Last_Modified` with field type "Last modified time" and select all columns you want to track changes to on a date/time basis moving forward.

## Record Views
1. Select **Kanban** from the Views options and group by `Trello_List` to see your tasks similar to how you saw them in Trello, complete with attachment covers! A _bonus_ is that if you have multiple images attached to a card, you can view them without expanding the card by just hovering over the attachment cover!
2. Move, collapse, and delete stacks as you see fit. Customize cards with as little or as much information as you want.

## Tasks vs Subtasks
There are probably many ways to creatively group your records to parallel the checklist option within a Trello card moving forward. 
- The most straightforward is to use the basic checklist formatting within the `Description` field to create lists
- Another is to think of your primary field `Tasks` instead as **'subtasks'** and create a new column to serve as the umbrella **'task'**. This new **'task'** column would be field type "single selection", then you could group your records by **'task'**.
